Data preparation

### 2024-02-12 procesamiento/01-prod_prep-micro.Rmd
a1

@@

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$digclass11d <- pmax(df1$digclass11, #social class respondent + spouse vs:
                     df1$digclass10spof,#social class spouse
                     na.rm = T) # if NA, do nothing
frq(df1$digclass11d)
df1$digclass11d <- 
  sjlabelled::set_labels(df1$digclass11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                         labels =  c("I", "II", "IIIab", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
sjlabelled::set_label(df1$digclass11d) <- "Social Class"
frq(df1$digclass11d)

@@

sjlabelled::set_label(df1$digclass3) <- "Social Class"
frq(df1$digclass3)

@@

df1$dclass11d <- 
  sjlabelled::set_labels(df1$dclass11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10,11),
                         labels =  c("I", "II", "IIIa","IIIb", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
frq(df1$dclass11d)
sjlabelled::set_label(df1$dclass11d) <- "Social Class"

# Class Scheme 3 classes
df1$dclass3 <- car::recode(df1$dclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV+V)';
                             8:11='Working Class (VI+VII)'")

sjlabelled::set_label(df1$dclass3) <- "Social Class"
frq(df1$dclass3)
frq(df1$class3)
frq(df1$digclass3)
```

@@

- using my original coding
```{r homogeneity-v01}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

# df1$eindex<- (df1$outgroup-df1$ingroup)/(df1$outgroup+df1$ingroup)
df1$eindex<- (df1$ingroup-df1$outgroup)/(df1$outgroup+df1$ingroup)
sjlabelled::set_label(df1$eindex) <- "EI homophily index"
#frq(df1$eindex)
#hist.default(df1$eindex)
df1$eindex_gc <- group_center(x = df1$eindex,grp = df1$country)
#hist.default(df1$eindex_gc)

df1 %>% select(class3,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass) <- "Class homogeneity"
df1$homclass_gc <- group_center(x = df1$homclass,grp = df1$country)
# View(df1[,c("homclass","homclass_gc")])
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

@@

@@

```{r homogeneity-v02}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(digclass3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

@@

```{r homogeneity-R-digclass}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

@@

sjmisc::frq(df1$week_cont)
sjPlot::tab_corr(df1[,c("know_total","week_cont","homclass")])

>>>>

df1$class3_V <- car::recode(df1$class11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$class3_V) <- "Social Class (V to Working Class)"
frq(df1$class3_V)

@@

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$digclass11d <- pmax(df1$digclass11, #social class respondent + spouse vs:
                     df1$digclass10spof,#social class spouse
                     na.rm = T) # if NA, do nothing
frq(df1$digclass11d)
df1$digclass11d <- 
  sjlabelled::set_labels(df1$digclass11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                         labels =  c("I", "II", "IIIab", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
sjlabelled::set_label(df1$digclass11d) <- "Social Class (isko - v2)"
frq(df1$digclass11d)

@@

sjlabelled::set_label(df1$digclass3) <- "Social Class (isko - v2) "
frq(df1$digclass3)

df1$digclass3_V <- car::recode(df1$digclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$digclass3_V) <- "Social Class (V to Working Class - isko - v2)"
frq(df1$digclass3_V)

@@

df1$dclass11d <- 
  sjlabelled::set_labels(df1$dclass11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10,11),
                         labels =  c("I", "II", "IIIa","IIIb", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
frq(df1$dclass11d)
sjlabelled::set_label(df1$dclass11d) <- "Social Class (EGP11 -  DIGICLASS)"

# Class Scheme 3 classes
df1$dclass3 <- car::recode(df1$dclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:8='Intermediate class (III+IV+V)';
                             9:11='Working Class (VI+VII)'")

sjlabelled::set_label(df1$dclass3) <- "Social Class (EGP3 -  DIGICLASS)"

# Class Scheme 3 classes -  V to working class
df1$dclass3_V <- car::recode(df1$dclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$dclass3_V) <- "Social Class (V to Working Class - DIGCLASS)"
frq(df1$dclass3_V)

# Class Scheme 3 classes -  IIIb to working class

df1$dclass3_III <- car::recode(df1$dclass11d,as.factor = T,
                            levels =c('Service Class',
                                      'Intermediate class',
                                      'Working Class'), 
                            "c(1,2)='Service Class';c(3,5,6,7,8)='Intermediate class';c(4,9,10,11)='Working Class'")

sjlabelled::set_label(df1$dclass3_III) <- "Social Class (IIIb to Working Class - DIGCLASS)"
frq(df1$dclass3_III)

# Class Scheme 3 classes -  IIIb and V to working class
df1$dclass3_III_V <- car::recode(df1$dclass11d,as.factor = T,
                            levels =c('Service Class',
                                      'Intermediate class',
                                      'Working Class'), 
                            "c(1,2)='Service Class';c(3,5,6,7)='Intermediate class';c(4,8,9,10,11)='Working Class'")

sjlabelled::set_label(df1$dclass3_III_V) <- "Social Class (IIIb+V to Working Class - DIGCLASS)"
frq(df1$dclass3_III)

# Class Scheme 6 classes
df1$dclass6 <- car::recode(
  df1$dclass11d,
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$dclass6) <- "Social Class (DIGCLASS)"
```

@@

- using my original coding
```{r homogeneity-v01-V-to-intermediate}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

# df1$eindex<- (df1$outgroup-df1$ingroup)/(df1$outgroup+df1$ingroup)
df1$eindex<- (df1$ingroup-df1$outgroup)/(df1$outgroup+df1$ingroup)
sjlabelled::set_label(df1$eindex) <- "EI homophily index (isko v1 - V to Intermediate)"
#frq(df1$eindex)
#hist.default(df1$eindex)
df1$eindex_gc <- group_center(x = df1$eindex,grp = df1$country)
#hist.default(df1$eindex_gc)

df1 %>% select(class3,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass) <- "Class homogeneity (V to Intermediate)"
df1$homclass_gc <- group_center(x = df1$homclass,grp = df1$country)
# View(df1[,c("homclass","homclass_gc")])
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

@@

```{r homogeneity-v01-V-to-working}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3_V,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3_V=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3_V=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3_V=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(class3_V,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$class3_V=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3_V=="Intermediate class (III+IV)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3_V=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

# df1$eindex<- (df1$outgroup-df1$ingroup)/(df1$outgroup+df1$ingroup)
df1$eindex_V<- (df1$ingroup-df1$outgroup)/(df1$outgroup+df1$ingroup)
sjlabelled::set_label(df1$eindex_V) <- "EI homophily index (isko v1 - V to Working)"
#frq(df1$eindex)
#hist.default(df1$eindex)
df1$eindex_gc <- group_center(x = df1$eindex,grp = df1$country)
#hist.default(df1$eindex_gc)

df1 %>% select(class3_V,n_high,n_middle,n_low,outgroup,ingroup,eindex_V)
df1$homclass_V <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass_V) <- "Class homogeneity (isko v1 - V to Working)"
df1$homclass_V_gc <- group_center(x = df1$homclass_V,grp = df1$country)
# View(df1[,c("homclass","homclass_gc")])
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

@@

```{r homogeneity-v02-V-intermediate}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(digclass3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

```{r homogeneity-v02-V-working}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(digclass3_V,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$digclass3_V=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3_V=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3_V=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(digclass3,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$digclass3_V=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$digclass3_V=="Intermediate class (III+IV)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$digclass3_V=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

# df1$eindex<- (df1$outgroup-df1$ingroup)/(df1$outgroup+df1$ingroup)
df1$eindex2_V<- (df1$ingroup-df1$outgroup)/(df1$outgroup+df1$ingroup)
sjlabelled::set_label(df1$eindex2_V) <- "EI homophily index (isko - v2- V to Working)"
#frq(df1$eindex)
#hist.default(df1$eindex2)
df1$eindex2_V_gc <- group_center(x = df1$eindex2_V,grp = df1$country)
#hist.default(df1$eindex2_gc)

df1 %>% select(dclass3_V,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass2_V <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass2_V) <- "Class homogeneity (isko - v2- V to Working)"
df1$homclass2_V_gc <- group_center(x = df1$homclass2_V,grp = df1$country)
# View(df1[,c("homclass","homclass_gc")])
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

@@

```{r homogeneity-R-digclass-V-intermediate}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

@@

```{r homogeneity-R-digclass-V-working}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(dclass3_V,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$dclass3_V=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3_V=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3_V=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(dclass3_V,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$dclass3_V=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3_V=="Intermediate class (III+IV)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3_V=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

# df1$eindex<- (df1$outgroup-df1$ingroup)/(df1$outgroup+df1$ingroup)
df1$eindex3_V<- (df1$ingroup-df1$outgroup)/(df1$outgroup+df1$ingroup)
sjlabelled::set_label(df1$eindex3_V) <- "EI homophily index"
#frq(df1$eindex)
#hist.default(df1$eindex)
df1$eindex3_V_gc <- group_center(x = df1$eindex3_V,grp = df1$country)
#hist.default(df1$eindex_gc)

df1 %>% select(dclass3,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass3_V <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass3_V) <- "Class homogeneity (DIGCLASS version - V to Working)"
df1$homclass3_V_gc <- group_center(x = df1$homclass3_V,grp = df1$country)
# View(df1[,c("homclass","homclass_gc")])
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

```{r homogeneity-R-digclass-IIIb-working}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(dclass3_III,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$dclass3_III=="Service Class",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3_III=="Intermediate class",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3_III=="Working Class",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(dclass3_III,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$dclass3_III=="Service Class",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3_III=="Intermediate class",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3_III=="Working Class",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

df1 %>% select(dclass3_III,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass3_III <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass3_III) <- "Class homogeneity (DIGCLASS version - IIIb to Working)"
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

```{r homogeneity-R-digclass-IIIb-V-working}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(dclass3_III_V,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$dclass3_III_V=="Service Class",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3_III_V=="Intermediate class",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$dclass3_III_V=="Working Class",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(dclass3_III_V,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$dclass3_III_V=="Service Class",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3_III_V=="Intermediate class",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$dclass3_III_V=="Working Class",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

df1 %>% select(dclass3_III_V,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass3_III_V <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass3_III_V) <- "Class homogeneity (DIGCLASS version - IIIb+V to Working)"
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

@@

sjmisc::frq(df1$week_cont)
sjPlot::tab_corr(df1[,c("know_total","week_cont","homclass","homclass_V","homclass2","homclass2_V")])

df1$homclass
